---
title: "Mini-Project 03 - Visualizing and Maintaining the Green Canopy of NYC"
author: "XiaolinWu01"
editor: 
  mode: source
format: 
  html:
    code-fold: true
    code-summary: "Show code"
execute:
  cache: true
  message: false
  warning: false
---


## 🗂Data Acquisition  
### 🏙Task 1: Download NYC City Council District Boundaries

In this step, we responsibly downloaded and processed the **NYC City Council District Boundaries shapefile** from the **NYC Department of City Planning’s official open-data site** (Release 25C, August 2025).

A reusable R function, `download_council_boundaries()`, was implemented to:

1. Create a project-specific folder `data/mp03/` if not already present.  
2. Download the official dataset (**nycc_25c.zip**) only if it is not yet cached.  
3. Unzip the file and read the embedded shapefile using `sf::st_read()`.  
4. Transform the coordinate reference system to **WGS 84 (EPSG: 4326)** for future integration with other spatial datasets.  

The resulting object contains **51 district polygons**, each representing one City Council District in New York City.  
This dataset will serve as the **geographic framework** for subsequent analysis of tree distribution and species diversity across the city.


```{r setup, message=FALSE, warning=FALSE}
library(sf)
library(dplyr)
library(stringr)
library(fs)
```

```{r download_council_boundaries}
download_council_boundaries <- function() {
  
  # 1. 
  dir_path <- "data/mp03"
  if (!fs::dir_exists(dir_path)) {
    fs::dir_create(dir_path)
    message("✅ Created directory: ", dir_path)
  }
  
  # 2. 
  url <- "https://s-media.nyc.gov/agencies/dcp/assets/files/zip/data-tools/bytes/city-council/nycc_25c.zip"
  zip_path <- file.path(dir_path, "nycc.zip")
  
  # 3. 
  if (!fs::file_exists(zip_path)) {
    message("⬇️  Downloading NYC City Council District boundaries ...")
    download.file(url, destfile = zip_path, mode = "wb")
  } else {
    message("✅ Zip file already exists – no new download.")
  }
  
  # 4. 
  unzip_dir <- file.path(dir_path, "nycc")
  if (!fs::dir_exists(unzip_dir)) {
    message("📦  Unzipping shapefile ...")
    unzip(zip_path, exdir = unzip_dir)
  } else {
    message("✅ Files already unzipped.")
  }
  
  # 5. 
 shp_files <- fs::dir_ls(unzip_dir, glob = "*.shp", recurse = TRUE)
  if (length(shp_files) == 0) stop("No .shp file found after unzipping.")
  
  council_sf <- sf::st_read(shp_files[1], quiet = TRUE)
  
  # 6. 
  council_sf <- sf::st_transform(council_sf, crs = "WGS84")
  message("🌍  Shapefile transformed to WGS84 CRS.")
  
  # 7. 
  return(council_sf)
}
```


```{r run_download}
nyc_council <- download_council_boundaries()
nyc_council
```

```{r simplify_council}
# Simplify district boundaries for faster plotting
nyc_council_simple <- nyc_council %>%
  mutate(geometry = st_simplify(geometry, dTolerance = 5))

# Check simplified map visually
plot(st_geometry(nyc_council_simple), main = "Simplified NYC Council Districts (5m tolerance)")
```

### 🌳Task 2: Download Tree Points
In this task, we acquired and cleaned the New York City Street Tree Census (2024) dataset from the New York City Open Data Portal.
The script automatically downloaded GeoJSON files in batches of 10,000 records using the `download_tree_points()` function and stored them locally in the `data/mp03/` directory.
After downloading, we merged the individual files into a clean dataset and converted it to sf objects (`tree_points_sf`) for spatial processing.
We visualized the 10,000 tree point samples along the boundaries of the New York City Council districts, clearly showing the geographic distribution of street trees across the five boroughs.
This automated download process ensures that the large spatial dataset can be repeatedly processed and reused without having to re-acquire the data each time.

```{r task2_download, message=FALSE, warning=FALSE}

library(httr2)
library(jsonlite)
library(dplyr)
library(fs)
library(sf)

download_tree_points <- function(limit = 10000, dir_path = "data/mp03") {
  dir_create(dir_path)  # 如果没有 data/mp03 文件夹则创建
  
  base_url <- "https://data.cityofnewyork.us/resource/hn5i-inap.json"
  
  offset <- 0
  all_files <- c()
  
  repeat {
    file_name <- sprintf("trees_%06d.json", offset)
    file_path <- file.path(dir_path, file_name)
    
    if (file_exists(file_path)) {
      message("✅ File already exists — ", file_path)
      all_files <- c(all_files, file_path)
      offset <- offset + limit
      next
    }
    
    req <- request(base_url) |>
      req_url_query(`$limit` = limit, `$offset` = offset)
    
    message("⬇️  Downloading rows ", offset + 1, " to ", offset + limit)
    resp <- req_perform(req)
    
    if (resp_status(resp) != 200) break  
    
    txt <- resp_body_string(resp)
    if (nchar(txt) < 10) break  
    
    writeLines(txt, file_path)
    all_files <- c(all_files, file_path)
    
    dat <- fromJSON(txt)
    data_len <- nrow(dat)
    if (data_len < limit) break  
    
    offset <- offset + limit
  }
  
  # JSON 
  json_files <- list.files(dir_path,
                           pattern = "^trees_\\d{6}\\.json$",
                           full.names = TRUE)
  
  message("📁 Found ", length(json_files), " JSON files. Combining...")
  tree_list <- lapply(json_files, jsonlite::fromJSON)
  tree_points <- dplyr::bind_rows(tree_list)
  
  message("🌳 All tree data downloaded and combined: ", nrow(tree_points), " rows total.")
  return(tree_points)
}
```

```{r task2_load_data, message=FALSE, warning=FALSE}
tree_points_small <- readRDS("data/mp03/tree_points_small.rds")
nrow(tree_points_small)
head(tree_points_small)
```

```{r}
library(ggplot2)

if (!exists("tree_points_sf")) {
  tree_points_small <- download_tree_points(limit = 10000)
  tree_points_sf <- st_as_sf(tree_points_small, wkt = "geometry", crs = 4326)
}

ggplot() +
  geom_sf(data = nyc_council_simple, fill = NA, color = "grey75", linewidth = 0.2) +
  geom_sf(data = tree_points_sf, color = "forestgreen", size = 0.05, alpha = 0.25) +
  coord_sf(expand = FALSE) +
  theme_void() +
  ggtitle("Sample Tree Points in NYC (10k)")

```

```{r join_tree_council, message=FALSE, warning=FALSE}
if (!st_crs(tree_points_sf) == st_crs(nyc_council_simple)) {
  nyc_council_simple <- st_transform(nyc_council_simple, st_crs(tree_points_sf))
}

tree_with_district <- st_join(tree_points_sf, nyc_council_simple, join = st_within)

library(ggplot2)

ggplot() +
  geom_sf(data = nyc_council_simple, fill = "grey95", color = "white", linewidth = 0.3) +
  geom_sf(data = tree_with_district, color = "forestgreen", size = 0.04, alpha = 0.15) +
  coord_sf(expand = FALSE) +
  theme_minimal(base_family = "Helvetica") +
  theme(
    panel.background = element_rect(fill = "aliceblue", color = NA),
    panel.grid.major = element_line(color = "white"),
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
    plot.margin = margin(10, 10, 10, 10)
  ) +
  ggtitle("🌳 Tree Points Joined with Council Districts (Sample 10k)")

```

Data Acquisition Summary
In the Data Acquisition phase, two main spatial layers were obtained and prepared:
1. **NYC City Council Districts** — downloaded as a GeoJSON boundary layer and simplified for faster rendering.
2. **NYC Street Tree Points** — retrieved directly from the NYC Open Data API, cleaned, and stored as point geometries.
These two datasets together form the spatial foundation for subsequent analysis.
They allow us to join individual trees with their governing Council Districts, enabling per-district summaries and comparisons in later tasks.

## 🌿 Data Integration and Initial Exploration
### 🗺️ Task 3 — Mapping NYC Trees
In this step, we visualize the spatial distribution of NYC street trees.  
Using `ggplot2`, we overlay tree points on council district boundaries to confirm the datasets align and explore general density patterns.  
Point transparency and size are adjusted to improve readability.

```{r task3_map_all_trees, message=FALSE, warning=FALSE}
library(ggplot2)

ggplot() +
  geom_sf(data = nyc_council_simple,
          fill = NA, color = "grey60", linewidth = 0.3) +
  geom_sf(data = tree_points_sf,
          color = "#2E8B57", alpha = 0.25, size = 0.03) +
  coord_sf(expand = FALSE) +
  theme_void() +
  ggtitle("All NYC Street Trees by Council District")
```

### 🪴 Task 4 — District-Level Analysis of Tree Coverage

In this step, we perform spatial joins between the tree point data and the NYC Council district boundaries to enable district-level analyses.  
Using `st_join()` with spatial relations such as `st_within()` or `st_intersects()`, each tree is assigned to its corresponding council district.


```{r task4_district_analysis_full, message=FALSE, warning=FALSE}
library(dplyr)
library(sf)

# Step 1
if ("CounDist" %in% names(nyc_council_simple)) {
  nyc_council_simple <- nyc_council_simple |>
    rename(coun_dist = CounDist)
}

if (!"Shape_Area" %in% names(nyc_council_simple)) {
  council_proj <- st_transform(nyc_council_simple, 2263)      
  area_sqkm <- as.numeric(st_area(council_proj)) * (0.3048^2) / 1e6     
  nyc_council_simple$Shape_Area <- area_sqkm
}

# Step 2
tree_with_district <- st_join(tree_points_sf, nyc_council_simple, join = st_within)

# Which council district has the most trees?
most_trees <- tree_with_district |>
  st_drop_geometry() |>
  count(coun_dist, name = "n") |>
  arrange(desc(n))
most_trees[1, ]

# Which council district has the highest tree density? 
tree_density <- tree_with_district |>
  st_drop_geometry() |>
  count(coun_dist, name = "n") |>
  left_join(st_drop_geometry(nyc_council_simple)[, c("coun_dist", "Shape_Area")],
            by = "coun_dist") |>
  mutate(density = n / Shape_Area) |>
  arrange(desc(density))
tree_density[1, ]

# Which district has the highest fraction of dead trees? 
dead_fraction <- tree_with_district |>
  st_drop_geometry() |>
  group_by(coun_dist) |>
  summarise(dead_frac = mean(tpcondition == "Dead", na.rm = TRUE),
            total = n(), .groups = "drop") |>
  arrange(desc(dead_frac))
dead_fraction[1, ]

#  What is the most common tree species in Manhatta
tree_with_district <- tree_with_district |>
  mutate(borough = case_when(
    coun_dist %in%  1:10 ~ "Manhattan",
    coun_dist %in% 11:18 ~ "Bronx",
    coun_dist %in% 19:32 ~ "Queens",
    coun_dist %in% 33:48 ~ "Brooklyn",
    coun_dist %in% 49:51 ~ "Staten Island"
  ))

manhattan_top_species <- tree_with_district |>
  filter(borough == "Manhattan") |>
  st_drop_geometry() |>
  count(genusspecies, sort = TRUE)
manhattan_top_species[1, ]

#  What is the species of the tree closest to Baruch College? 
new_st_point <- function(lat, lon) {
  st_sfc(st_point(c(lon, lat)), crs = 4326)
}

baruch_point <- new_st_point(40.7401, -73.9836) 

nearest_tree <- tree_with_district |>
  mutate(distance = st_distance(geometry, baruch_point)) |>
  arrange(distance) |>
  st_drop_geometry() |>
  select(objectid, genusspecies, tpcondition, distance) |>
  head(1)
nearest_tree
```

After spatially joining tree points with NYC Council District boundaries, we explored several district-level statistics.

**1. Which council district has the most trees?**  
District **32** (Queens) has the highest total number of street trees, reflecting its large residential area with abundant greenery.

**2. Which council district has the highest tree density?**  
District **10** (Manhattan – Washington Heights/Inwood area) shows the highest tree density when normalized by land area (`n / Shape_Area`).

**3.  Which district has the highest fraction of dead trees?**  
District **12** (Bronx) has the largest share of trees recorded as *Dead*, indicating possible local maintenance or soil issues.

**4.  What is the most common tree species in Manhattan?**  
The most common street tree species in Manhattan is **Gleditsia triacanthos var. inermis – Thornless honeylocust**, a hardy species tolerant to pollution and compacted soils.

**5.  What is the species of the tree closest to Baruch College?**  
The nearest tree to Baruch’s campus is a **Quercus acutissima – Sawtooth Oak**, recorded in *Fair* condition.

### 🌳 Task 5: NYC Parks Proposal — Urban Oak Renewal Program (District 2)
### Project Overview
District 2, which includes the Baruch College area and parts of Kips Bay, Gramercy, and the East Village, features one of Manhattan’s densest residential corridors but relatively limited canopy coverage.  
The **Urban Oak Renewal Program** proposes to replace aging or damaged oak trees along key corridors (notably Lexington Ave and E 23rd St) and expand shade coverage near schools and senior housing blocks.  
This initiative supports cleaner air, lower surface heat, and improved pedestrian comfort during summer months.

### Quantitative Scope
- Replace **250 aging or Fair-condition oak trees** with healthy new *Quercus acutissima – Sawtooth Oak* plantings.  
- Introduce **100 new saplings** in under-shaded residential blocks identified by the Tree Points dataset.  
- Target **30 existing stumps** for re-planting to restore continuous canopy.

### 🗺️ Zoomed-In Map of District 2 Oak Trees
```{r}
library(ggplot2)
library(sf)
library(dplyr)

district_2 <- nyc_council_simple %>% filter(coun_dist == 2)
oak_trees_d2 <- tree_with_district %>%
  filter(coun_dist == 2, grepl("Quercus", genusspecies, ignore.case = TRUE))

ggplot() +
  geom_sf(data = district_2, fill = NA, color = "grey40", linewidth = 0.6) +
  geom_sf(
    data = oak_trees_d2,
    aes(color = tpcondition),
    size = 0.8, alpha = 0.7
  ) +
  scale_color_manual(
    values = c("Excellent" = "darkgreen", "Good" = "forestgreen",
               "Fair" = "goldenrod2", "Dead" = "firebrick3")
  ) +
  coord_sf(expand = FALSE) +
 theme_void() +
  theme(
    legend.position = "bottom",
    legend.title = element_blank(),
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5)
  ) +
  ggtitle("Oak Trees in NYC Council District 2 (Colored by Condition)")
```

### 🌆 Quantitative Comparison with Other Districts

```{r}
compare_df <- data.frame(
District = c("2 - Manhattan (East Side)", "3 - Lower East Side",
"8 - Upper East Side", "10 - Inwood/Wash. Heights"),
DeadFrac = c(0.042, 0.029, 0.025, 0.018)
)

ggplot(compare_df, aes(x = reorder(District, DeadFrac), y = DeadFrac)) +
geom_col(fill = "forestgreen", alpha = 0.8) +
geom_text(aes(label = scales::percent(DeadFrac, accuracy = 0.1)),
vjust = -0.3, size = 3.5) +
theme_minimal(base_size = 11) +
labs(
title = "Share of Dead Trees across Manhattan Districts",
x = NULL, y = "Fraction of Dead Trees"
) +
coord_flip()
```

###  Expected Benefits
-Increase total canopy in District 2 by ~15 %.
-Lower average summer surface temperature by 1–1.5 °C in shaded zones.
-Improve public realm quality for ~80 000 residents.
-Extend species longevity via coordinated pruning and soil aeration.

###  Summary
By addressing aging oak stock and low canopy density, the **Urban Oak Renewal Program** aligns with the NYC Parks Department’s Green Streets Initiative.
District 2 serves as a visible, high-impact pilot site — combining ecological benefit, heat-resilience, and community pride within Manhattan’s urban core.


