---
title: "Xiaolin Wu — STA 9750 Website"
---

# Welcome!

Hi, my name is Xiaolin Wu.\
This website hosts my submission materials for **STA 9750 at Baruch College**.

I am currently a graduate student with interests in data analytics and urban design.\
Here, you’ll find my course projects, assignments, and reflections.

-   📄 [My Resume](Xiaolin_Wu_Resume.pdf)\
-   📧 Contact: xiaolin.wu1\@baruchmail.cuny.edu

```{r}
#| echo: false
#| message: false
#| warning: false

if(!require("leaflet")){
    options(repos=c(CRAN="https://cloud.r-project.org"))
    install.packages("leaflet")
    stopifnot(require("leaflet"))
}

baruch_longitude <- -73.98333
baruch_latitude  <- +40.75028

leaflet() |>
  addTiles() |>
  setView(baruch_longitude, baruch_latitude, zoom=17) |>
  addPopups(baruch_longitude, baruch_latitude, 
            "I am a Master's student at <b>Baruch College</b>!")
```

```{r}
#| include: false
1+1
```
--------------
Last Updated: `r format(Sys.time(), "%A %m %d, %Y at %H:%M%p")`

